---
title: " An Introduction to Geocoding with 'sf'"
author: "Giulia Maria Petrilli, Julia Wu and Jennifer Estigene"
date: '2023-10-30'
output: 
 html_document:
    toc: TRUE
    df_print: paged
    number_sections: FALSE
    highlight: monochrome
    theme: yeti
    toc_depth: 3
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(sf)
library(tmap)
#library(tm)
library(tmaptools)
library(stringr)
library(rvest)
library(geosphere)
library(nominatim)

library(ggmap)
library(mapview)

```


## Transforming point data to a sf object

We are from Rome and Tianjin. Let's see where these are on a map 

```{r}

cities <- data.frame(
  place = c("Rome", "Tianjin", "Ottawa"),
  long = c(12.496366, 117.190182, -75.690308),  
  lat = c(41.902782, 39.084158, 45.421529),    
  value = c(500, 600, 700))                     

class(cities)  # WHY DO WE CHECK CLASS

cities_dsf <- st_as_sf(cities, coords = c("long", "lat")) # st_as_sf function is used to convert data, such as a data frame or a matrix, into an sf object. An sf object is a special data structure used for working with spatial data. 
st_crs(cities_dsf) <- 4326
class(cities_dsf)

mapview(cities_dsf)

```

Quite cool! But to insert the coordinates feels a bit lenghty, especially when there is a r function that does it for you/ I am referring to geocode_OSM() 
```{r}

cities_c <- c("Rome", "Tianjin", "Ottawa")
geocoded_cities <- geocode_OSM(cities_c) 

geocoded_cities

# Create new column using st_as_sf to store the geometry in a column 

geometry_geocoded_cities <- st_as_sf(geocoded_cities,
                      coords = c(x = "lon", y = "lat"),
                      crs = 4326)

geometry_geocoded_cities
mapview(geometry_geocoded_cities)


```


Pretty cool because when only did geocode("Rome"), you are required to use an API address, so going to google and scrape that.  Just for demostration, this is what happens.  

``` {r}
geocode_OSM("Rome")

```

## North Carolina 

```{r}

demo(nc, ask = FALSE, echo = FALSE)
plot(st_geometry(nc))
# https://r-spatial.github.io/sf/articles/sf5.html

```
Color key place and size
In case a single attribute is selected, by default a color key is given the side of the plot where it leaves as much as possible room for the plotted map; for nc this is below:


```{r}
plot(nc["AREA"])
# https://r-spatial.github.io/sf/articles/sf5.html
```
Let's use st_centroid to find out about the centroids in each region in Corth carolina. Centroids tend to be the areas of easiest reach. For urban planning, infrastructure development, and resource allocation, knowing the centroids of regions can help authorities decide where to build facilities such as roads, schools, hospitals. Of course, it depends on the specific region, 


```{r}


plot(st_geometry(nc), border = 'grey', axes = TRUE) 
plot(st_geometry(st_centroid(nc)), pch = 3, col = 'red', add = TRUE)


# https://r-spatial.github.io/sf/articles/sf5.html

```
 This code space here is for general trials 
```{r}


```

## Excercise 3
## Excercise 4
## Excercise 5

